Noise-Protected Quantum Memory Registers with Nitrogen-Vacancy Centers and $^{13}$C Nuclear Spins

### 2016-06-04 swap-nvst-paper/paper.tex
add note about reading out the state of the ST qubit
added \fixme command to paper

%%% markup
\usepackage{color,soul} % text color and other editing options
% \ul{underline}, \st{strikethrough}, and \hl{highlight}
\usepackage[draft]{fixme} % in text notes
% \fxsetup{layout=footnote}
\newcommand{\nt}{\fxnote}
\newcommand{\re}[1]{{\bf \color{red} #1}}
\newcommand{\bl}[1]{{\bf \color{blue} #1}}

@@

Nitrogen-vacancy (NV) centers in diamond have been recognized in
recent years as a promising platform for quantum sensing and
information technologies. The electronic ground and optically excited
states of NV$^-$ split into spin triplets which can be polarized,
detected, and coherently manipulated with high fidelity with fields
and microwave radiation \cite{dobrovitski2013quantum}. Hyperfine
coupling between the NV center electron and $^{13}$C nuclei in bulk
diamond allows for the coherent control and measurement of individual
nuclear spins, and the realization of universal set of high-fidelity
quantum gates \cite{dobrovitski2013quantum, casanova2016noise}. These
properties have prompted extensive studies of NV centers for quantum
metrology and information processing applications
\cite{mamin2013nanoscale, steinert2010high, wang2015positioning,
  chou2015optimal, childress2006fault, yao2012scalable}. % fixme: the
% first and last sentences of this paragraph are essentially the same

@@

In this paper, we propose using the singlet-triplet subspace of two
$^{13}$C nuclear spins with identical effective larmor frequencies as
a logical qubit for storing quantum memory. Here identical effective
larmor frequencies implies that, absent of symmetry-breaking control
fields, the nuclei are indistinguishable from the perspective of the
NV electron. This indistinguishability protects the logical qubit from
dephasing via coupling to the NV electron spin, potentially increasing
its coherence time by a few orders of magnitude. % fixme: elaborate
We investigate the efficacy of this quantum memory register when
realized via the recently developed dynamical decoupling protocol
dubbed adaptive XY-$n$ (AXY-$n$) \cite{casanova2015robust}.

@@

When employing a dynamic decoupling protocol on resonance with an
effective larmor frequency $\omega_r$, one can apply an additional
magnetic decoupling field
$\v B_\t{dec}^{\p\omega}\p{t}=\v B_\t{dec}\cos\p{\omega_r
  t-\phi_\t{dec}}$ and realize, with a rotating wave approximation in
the frame of $H_\t{nZ}^\t{eff}$, the Hamiltonian
\begin{align}
  \tilde H_\t{int}^\t{dec} = \sum_{j:\omega_j=\omega_r}
  \tilde h_j^\t{int}\sigma_z^{NV}\v\sigma^j\c\tilde{\v n}_j,
  \label{H_int_dec}
\end{align}
where
\begin{align}
  \tilde h_j^\t{int}
  = h_j^\t{int}\p{\uv A_j^\perp\c\uv B_{\t{dec},j}^\perp}
  = h_j^\t{int}A_{j,\t{dec}}^\perp/A_j^\perp
\end{align}
and
\begin{align}
  \tilde{\v n}_j
  &= \uv A_{j,\t{dec}}^\perp\cos\phi_{DD}
    + \uv A_{j,\t{dec}}^{\perp\perp}\sin\phi_{DD}
\end{align}
with
\begin{align}
  \v A_{j,\t{dec}}^\perp
  &= \p{\v A_j^{\perp}\c\uv B_{\t{dec},j}^\perp}\uv B_{\t{dec},j}^\perp, \\
  \v A_{j,\t{dec}}^{\perp\perp}
  &= \uv\omega_j\times\v A_{j,\t{dec}}^\perp, \\
  \v B_{\t{dec},j}^\perp
  &= \v B_\t{dec} - \p{\v B_\t{dec}\c\uv\omega_j}\uv\omega_j.
\end{align} % fixme: provide derivation of these results?
In words: the decoupling field $\v B_\t{dec}^{\p\omega}$ effectively
suppresses the components of $\v A_j^\perp$ perpendicular to
$\uv B_\t{dec}^\perp$. These results to not depend on $\phi_\t{dec}$,
but do require that $A_j\ll\gamma_jB_{\t{dec},j}^\perp\ll\gamma_jB$.

@@

>>>>

%%% markup
\usepackage{color,soul} % text color and other editing options
% \ul{underline}, \st{strikethrough}, and \hl{highlight}
\newcommand{\fixme}[1]{{\bf \color{red} fixme: #1}}

@@

Nitrogen-vacancy (NV) centers in diamond have been recognized in
recent years as a promising platform for quantum sensing and
information technologies. The electronic ground and optically excited
states of NV$^-$ split into spin triplets which can be polarized,
detected, and coherently manipulated with high fidelity with fields
and microwave radiation \cite{dobrovitski2013quantum}. Hyperfine
coupling between the NV center electron and $^{13}$C nuclei in bulk
diamond allows for the coherent control and measurement of individual
nuclear spins, and the realization of universal set of high-fidelity
quantum gates \cite{dobrovitski2013quantum, casanova2016noise}. These
properties have prompted extensive studies of NV centers for quantum
metrology and information processing applications
\cite{mamin2013nanoscale, steinert2010high, wang2015positioning,
  chou2015optimal, childress2006fault, yao2012scalable}. \fixme{the
  first and last sentences of this paragraph are essentially the same}

@@

In this paper, we propose using the singlet-triplet subspace of two
$^{13}$C nuclear spins with identical effective larmor frequencies as
a logical qubit for storing quantum memory. Here identical effective
larmor frequencies implies that, absent of symmetry-breaking control
fields, the nuclei are indistinguishable from the perspective of the
NV electron. This indistinguishability protects the logical qubit from
dephasing via coupling to the NV electron spin, potentially increasing
its coherence time by a few orders of magnitude. \fixme{elaborate} We
investigate the efficacy of this quantum memory register when realized
via the recently developed dynamical decoupling protocol dubbed
adaptive XY-$n$ (AXY-$n$) \cite{casanova2015robust}.

@@

When employing a dynamic decoupling protocol on resonance with an
effective larmor frequency $\omega_r$, one can apply an additional
magnetic decoupling field
$\v B_\t{dec}^{\p\omega}\p{t}=\v B_\t{dec}\cos\p{\omega_r
  t-\phi_\t{dec}}$ and realize, with a rotating wave approximation in
the frame of $H_\t{nZ}^\t{eff}$, the Hamiltonian
\begin{align}
  \tilde H_\t{int}^\t{dec} = \sum_{j:\omega_j=\omega_r}
  \tilde h_j^\t{int}\sigma_z^{NV}\v\sigma^j\c\tilde{\v n}_j,
  \label{H_int_dec}
\end{align}
where
\begin{align}
  \tilde h_j^\t{int}
  = h_j^\t{int}\p{\uv A_j^\perp\c\uv B_{\t{dec},j}^\perp}
  = h_j^\t{int}A_{j,\t{dec}}^\perp/A_j^\perp
\end{align}
and
\begin{align}
  \tilde{\v n}_j
  &= \uv A_{j,\t{dec}}^\perp\cos\phi_{DD}
    + \uv A_{j,\t{dec}}^{\perp\perp}\sin\phi_{DD}
\end{align}
with
\begin{align}
  \v A_{j,\t{dec}}^\perp
  &= \p{\v A_j^{\perp}\c\uv B_{\t{dec},j}^\perp}\uv B_{\t{dec},j}^\perp, \\
  \v A_{j,\t{dec}}^{\perp\perp}
  &= \uv\omega_j\times\v A_{j,\t{dec}}^\perp, \\
  \v B_{\t{dec},j}^\perp
  &= \v B_\t{dec} - \p{\v B_\t{dec}\c\uv\omega_j}\uv\omega_j.
\end{align} \fixme{derive these results explicitly?}
In words: the decoupling field $\v B_\t{dec}^{\p\omega}$ effectively
suppresses the components of $\v A_j^\perp$ perpendicular to
$\uv B_\t{dec}^\perp$. These results to not depend on $\phi_\t{dec}$,
but do require that $A_j\ll\gamma_jB_{\t{dec},j}^\perp\ll\gamma_jB$.

@@

\fixme{when we wish to read out the ST qubit, it is possible to do so
  without performing the full coherent SWAP operation so long as we no
  longer need the larmor pair as a resource for the time being. Talk
  about this.}
